---
title: "ETC5521 Assignment 1"
author:
- Aphiaut Imuan
- Xintong You
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    fig_caption: yes
team: lorikeet
subtitle: Infrastructural Investment in USA
---

[This assignment is for ETC5521 Assignment 1 by Team `r rmarkdown::metadata$team` comprising of `r knitr::combine_words(rmarkdown::metadata$author)`.]{style="color:#006DAE;"}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE, 
                      warning = FALSE)
options(tinytex.verbose = TRUE)
library(tidyverse)
library(ggplot2)
library(knitr)
library(lubridate)
library(sf)
library(kableExtra)
library(broom)
library(readxl)
library(gridExtra)
library(bookdown)
library(tinytex)
```

```{r, echo=FALSE}
investment <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-08-10/investment.csv')
chain_investment <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-08-10/chain_investment.csv')
ipd <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-08-10/ipd.csv')
gdplev <- read_excel("gdplev.xlsx", range = "A9:C101", 
    col_names = FALSE, col_types = c("numeric", 
        "numeric", "numeric"))
gdplev <- gdplev %>%
  rename("year" = ...1,
         "GDP_current" = ...2,
         "GDP_chain" = ...3)
```


# Introduction and motivation


Infrastructure investment is an important part of the country. It includes federal investments such as water suppliers and private sector investments that manages electricity. It also includes the cost of research and development in technology.

Moreover, infrastructure investment plays a role in supporting both businesses and households. For example, the development of logistics in business sector has resulted in reduction of transportation cost. Households can access infrastructure services and can choose what and how to use. It also has a positive affect on both short-term and long-term economic growth. [(Stupak, 2017)](https://ecommons.cornell.edu/bitstream/handle/1813/78289/CRS_Economic_Impact_of_Infrastructure_Investment_0717.pdf?sequence=1&isAllowed=y)

As mentioned above, the case study should be of USA's infrastructure because USA is a big country. It has large population, and people can access infrastructure services easily. Furthermore, USA also has the world's largest economy. 

# Data description


The data comes from [Bureau of Economic Analysis](https://www.bea.gov/system/files/papers/BEA-WP2020-12.pdf). The raw .xlsx file is included or it can be downloaded directly from [BEA Working paper series](https://www.bea.gov/system/files/2021-01/infrastructure-data-may-2020.xlsx).

3 primary data sets are already cleaned and saved as .csv files and they all have five variables. Four out the variables are described below:

```{r, echo=FALSE}
variable <- c("category","meta_cat","group_num","year")
class <- c("character","character","double","double")
description <- c("Category of investment",
                 "Group category of investment",
                 "Group number of investment",
                 "Year of investment")
e.g. <- c("Private,Federal...","Total infrastructure,Water supply,GDP...","1,2,3...","1947,1948,1949...")
dic<- data.frame(variable, class,description, e.g.)

kable(dic,
      align = "llll","pipe")
```

The 3 data sets record gross investment, chained investment (USD in millions) and Implicit Price Deflators (IPDs) respectively.

```{r, echo=FALSE}
variable <- c("gross_inv","gross_inv_chain","gross_inv_ipd")
class <- c("double","double","double")
description <- c("Gross investment in millions of USD",
                 "Gross investment (chained 2021 dollars) in millions of USD",
                 "Implicit Price Deflators (IPDs) An implicit price deflator is the ratio of the current-dollar value of a series, such as gross domestic product (GDP), to its corresponding chained-dollar value, multiplied by 100.")
dic2 <- data.frame(variable,class,description) 

kable(dic2,
      align = "lll","pipe")

```

The gdplev data set is downloaded from [Bureau of Economic Analysis](https://www.bea.gov/data/gdp/gross-domestic-product), which is a supplementary data set for our analysis. It records the current GDP and chain GDP in US from 1929 to 2021. Vaariables in this dataset are *year*, *GDP_current* and *GDP_chain* whose datatype is double.

- Chained dollars: Chained dollars is a method of adjusting real dollar amounts for inflation over time, to allow the comparison of figures from different years. The U.S. Department of Commerce introduced the chained-dollar measure in 1996.

# Questions of interest


1. What category of investment is the highest average gross investment and Chain? Does it same?

2. What is the trend of total basic infrastructure?

3. what is the trend of total social infrastructure?

4. What year and category are the highest and lowest investment and Chain?

5. **What is the relationship between total digital investment and GDP?** 

6. What are the highest investment between Air, Water, Rail transportation investment ?

7. What are the similarities/difference of trend between Federal electric power structures investment and Private electric power structures investment?

8. How does the different type of investment change? (digital, transportation and power)

9. What is the proportion of total social investment in all category infrastructure?

10. What is the proportion of total basic infrastructure investment in all category infrastructure?

11. **What is the relationship between transportation investment (chained US dollars) and GDP (chained US dollars)?**

12. What are the trends in all categories infrastructure investment so far?

13. In 2012, what is the proportion of various investments in the total investment？

14. What category of investment has the greatest impact on GDP and how?

15. **What is the relationship between GDP and private investment, federal investment？And which one is stronger impact on GDP?**

16. What is the proportion of total digital investment in all category infrastructure?

17. **Is there a linear relationship between GDP and total basic infrastructure investment? Positive or negative?**

18. How each category is changed under Transportation from 2000 to 2017?

19. What is the total gross investment ipd for conservation and development in each category?

20. What is the relationship between current GDP and chained GDP?

# Expected findings


In question 5: We expect to find the positive relationship between total digital investment and GDP. Moreover we expect to see the steep line from 1900 to 2017.

In question 11: We expect the positive relationship between transportation (chained US dollar) and GDP (chained US dollar). This expected finding can be explained by linear model and there will be high effect value to GDP.

In question 15: We expect to find a strong relationship between GDP and federal investment and there is a positive relation both sides.

In question 17: We expect to find a positive linear relationship between GDP and total basic infrastructure investment.

# Analysis and findings

**Methodology**

This study uses the linear model to analyze because linear models are simple and present a mathematical equation which is easy to interpret and can make predictions. Moreover, linear regression model is a reliable predictor because it is a long established statistical procedure  and the character of linear regression is easy to understand. (IBM, 2022).


**What is the relationship between total digital investment and GDP?**

```{r}
investment <- investment %>%
  left_join(gdplev, by = "year") 
investment <- investment %>%
  mutate(GDP_current = GDP_current*1000)


chain_investment <- chain_investment %>%
  left_join(gdplev, by = "year")

chain_investment <- chain_investment %>%
  mutate(GDP_current = GDP_current*1000)
```

```{r q6fig1, fig.cap= "The trend of total digital infrastructure investment from 1947 to 2017", fig.align='center'}
q6 <- investment %>%
  filter(category == "Total digital infrastructure") %>%
  rename("digital" = gross_inv)
  
q6_lm <-  lm(GDP_current ~ digital, data = q6)
  
ggplot(q6,
       aes(x = year,
           y = digital)) +
  geom_point() +
  geom_vline(xintercept = 1990, linetype="dashed", color = "red")+
  theme_bw() +
  ggtitle("The trend of total digital infrastructure investment from 1947 to 2017")
```


```{r q6fig2, fig.cap= "The trend of Gross Domestic Product (GDP) from 1947 to 2017",fig.align='center'}
ggplot(q6,
       aes(x = year,
           y = GDP_current)) +
  geom_point() +
  geom_vline(xintercept = 1990, linetype="dashed", color = "red")+
  theme_bw() +
  ggtitle("The trend of Gross Domestic Product (GDP) from 1947 to 2017")
```

```{r q6fig3, fig.cap= "The relationship between total digital investment and gdp",fig.align='center'}
 ggplot(q6,
       aes(x = digital,
           y = GDP_current)) +
  geom_point(shape = 1,
               size = 2,
               color="#0072B2") +
  geom_smooth(method = "lm",
              col = "red",
              size = 1)+
  ggtitle("The relationship between total digital investment and gdp") +
   theme_bw()
 
 tidy(q6_lm) %>% kable() %>% 
  kable_classic(full_width = F, html_font = "Cambria")
q6_lm_summary <- glance(q6_lm)
q6_lm_summary %>% kable() %>% 
  kable_classic(full_width = F, html_font = "Cambria")

```
Figure \@ref(fig:q6fig1) shows the positive trend of total digital infrastructure investment. Moreover, it has a steeply increasing trend from 1990 to 2017. Overall there is an increasing trend but around 1999 to 2009 there is a slight fluctuation. Figure \@ref(fig:q6fig3) illustrates the proportional increase between Total digital infrastructure and GDP which is 1:1. 

The table of results can be expressed in the formula as:

$$
\begin{align*}
{GDP} = 659595.97 + 112.63{digital}
\end{align*}
$$

This formula means if total digital infrastructure increases by 1 million US dollars, the Gross Domestic Product will increase by 112.63 million US dollars. This result is related to the study of Zhang et al. (2022) who investigated that the increase of digital economy will increase the GDP by around 0.78%. Furthermore, he told that the increase is due to the development of new technology such as internet and mobile phone communication. R-squared is 95.54% of the variance and it shows a nice linear model. Therefore, the result of a relationship between total digital infrastructure and GDP can be summarized by the linear model.

In Addition, this resulted same as our expected finding which was a positive relationship between total digital infrastructure and GDP. However, the trend of Total digital infrastructure has steeply increased since 1990 and GDP's has steeply increased since 1980. It means total digital infrastructure doesn't significantly affect GDP from 1980 to 1990.


**What is the relationship between transportation investment (chained US dollars) and GDP (chained US dollars)?**

```{r q14fig4,fig.cap="The trend of transpotation investment and Gross Domestic Product (GDP) from 1947 to 2017",fig.align='center'}

q14 <- chain_investment %>%
   filter(category == "Transportation") %>%
   pivot_wider(names_from = category, values_from = gross_inv_chain)

q14_lm <-  lm(GDP_chain ~ Transportation, data = q14)
  
transport <- ggplot(q14,
       aes(x = year,
           y = Transportation)) +
  geom_point() +
  theme_bw() +
  ggtitle("The trend of transpotation investment from 1947 to 2017")
 # geom_vline(xintercept = 1990, linetype="dashed", color = "red")+
#  ggtitle("The trend of total digital infrastructure investment from 1947 to 2017")

gdp <- ggplot(q14,
       aes(x = year,
           y = GDP_chain)) +
  geom_point() +
  geom_vline(xintercept = 1990, linetype="dashed", color = "red")+
  theme_bw() +
  ggtitle("The trend of Gross Domestic Product (GDP) from 1947 to 2017")

grid.arrange(transport, gdp, nrow=2)
```

```{r q14fig5,fig.cap="The relationship between transportation investment and gdp",fig.align='center'}
 ggplot(q14,
       aes(x = Transportation,
           y = GDP_chain)) +
  geom_point(shape = 1,
               size = 2,
               color="#0072B2") +
  geom_smooth(method = "lm",
              col = "red",
              size = 1)+
  ggtitle("The relationship between transportation investment and gdp") +
   theme_bw()
 
 tidy(q14_lm)%>% kable() %>% 
  kable_classic(full_width = F, html_font = "Cambria")
q14_lm_summary <- glance(q14_lm)
q14_lm_summary%>% kable() %>% 
  kable_classic(full_width = F, html_font = "Cambria")

```

Figure \@ref(fig:q14fig4) shows the positive trend of transportation investment (chained dollar), however, there is a slight steep fluctuation between 1968 and 1998. While Gross Domestic Product (chained dollar) has a positive trend and increases over the years. Figure \@ref(fig:q14fig5) illustrates the relationship between transportation investment (chained dollar) and GDP (chained dollar),  nonetheless, it is difficult to explain the relationship between both variables as linear. 

The table of results can be expressed in the formula as:

$$
\begin{align*}
{GDP} = -5164.83 + 0.13{Transportation}
\end{align*}
$$

This formula means if transportation investment increases by 1 million chained US dollars, the Gross Domestic Product will increase by 0.13 million chained US dollars. This result relates to the study of Weisbrod and Reno (2009) that argued the positive significance of the relationship between transportation investment and GDP. Moreover, transportation investment has a positive effect on economic growth (Lin, 2020). R-squared is 67.39% of the variance. Therefore, the result of the relationship between transportation investment (chained dollar) and Gross Domestic Product (chained dollar) should not summarize by a linear model or this formula has omitted variables. 

This result same as our expected finding that is a positive relationship between transportation investment (chained dollar) and Gross Domestic Product (chained dollar). However, this relationship can not explain by the linear model and the coefficient is too low which is not the same as our expected finding.

**What is the relationship between GDP and private investment, federal investment？And which one is stronger impact on GDP?**

```{r q18fig6,fig.cap="The trends of private, federal investment and current GDP",fig.align='center'}
options(scipen=100, digits=4)
q18_P1 <- investment %>%
  filter(category == "Private") 
q18_P <- q18_P1 %>%
  filter(meta_cat == "Total infrastructure")
q18_F1 <- investment %>%
  filter(category == "Federal")
q18_F <- q18_F1 %>%
  filter(meta_cat == "Total infrastructure")

gg1 <- ggplot(q18_P,aes(year,GDP_current)) +
  geom_line(color = "blue") +
  theme_bw() +
  ggtitle("Trend of current GDP") 

gg2 <- ggplot(q18_P,aes(year,gross_inv)) +
  geom_line(color="blue") +
  theme_bw()+
  ggtitle("Trend of gross private investment") 

gg3 <-  ggplot(q18_F,aes(year,gross_inv)) +
  geom_line(color="blue") +
  theme_bw() +
  ggtitle("Trend of gross federal investment") 
  
grid.arrange(gg1,gg2,gg3,layout_matrix=rbind(c(1,1),c(2,3)))
```

```{r q18fig7, fig.cap="The relationship between investment and GDP",fig.align='center'}
ggplotRegression <- function (x) {
  ggplot(x$model, aes_string(x = names(x$model)[2], 
                               y = names(x$model)[1])) +
    geom_point(shape = 1,
               size = 2,
               color="#0072B2")+
    geom_smooth(method = "lm", 
                col = "red",
                size = 1) +
    labs(title = paste("R2 = ",signif(summary(x)$r.squared, 3)))
}

g1 <- ggplotRegression(lm(GDP_current ~ gross_inv,data = q18_P)) +
  labs(x = "Gross Private Investment", y = "Current GDP") +
  theme_bw()
g2 <- ggplotRegression(lm(GDP_current ~ gross_inv,data = q18_F)) +
  labs(x = "Gross Federal Investment", y = "Current GDP") +
  theme_bw()

grid.arrange(g1,g2,nrow = 1)
```

Figure \@ref(fig:q18fig6) illustrates that the current GDP is generally increasing over time, and the trend of gross private investment is similar to that. While the gross federal investment is fluctuating since 1980 Figure 9.

By observing the trends of the three variables, this study predicts the current GDP and the other two might have a linear relationship, then we use the linear model to judge whether they have a linear relationship. As a result, GDP does have a linear relationship with private investment and federal investment respectively, and private investment has a stronger impact on current GDP.

R-squared is an important statistical measure that represents the proportion of the variance for current GDP that is explained by the gross private and federal investment in the regression model. For example, Figure \@ref(fig:q18fig7) shows that 98.6% of private investment can explain the current GDP, rather only 78.1% of federal investment can explain that. Obviously, there is a stronger linear relationship between GDP and private investment which is unexpected. This result relates to the study of Private investment as the engine of economic growth and social welfare that investigated the private investment can explain the variance of GDP growth rate more than public investment (Doménech & Sicilia, 2021).

Therefore, this results same as our expected finding that is the relationship between both variable and GDP is a linear model. However, the R-squared of federal investment is not as expected finding because the R-squared of private investment is much more valuable.


**Is there a linear relationship between GDP and total basic infrastructure investment? Positive or negative?**

```{r q20fig8, fig.cap = "The relationship between GDP and total basic infrastructure investment",fig.align='center'}
q20 <- investment %>%
  filter(category == "Total basic infrastructure")

q20_lm <-lm(GDP_current ~ gross_inv, data = q20)

tidy(q20_lm)%>% kable() %>% 
  kable_classic(full_width = F, html_font = "Cambria")
q20_lm_summary <- glance(q20_lm)
q20_lm_summary%>% kable() %>% 
  kable_classic(full_width = F, html_font = "Cambria")


g3 <- ggplotRegression(lm(GDP_current ~ gross_inv,data = q20)) +
  labs(x = "Total Basic Infrastructure Investment", y = "Current GDP") +
  theme_bw() +
  ggtitle("Relationship between GDP and total basic infrastructure investment")
g3
```

From the result can be expressed the formula as:

$$
\begin{align*}
{GDP} = -337073.45	 + 59.02 {Basic}
\end{align*}
$$

This formula means if transportation investment increases by 1 million US dollars, the Gross Domestic Product will increase by 59.02  million US dollars. Moreover, the distribution of data has a positive linear relationship in figure \@ref(fig:q20fig8). The trend is increasing because the infrastructure investment effect increases a huge stock of public capital such as developing or creating new roads. That is a direct effect to increase government spending (G) in the GDP formula (Stupak, 2017).  Furthermore, Gunnion (2021) claims that investment in infrastructure has a positive impact on economic output. R-squared is 98.67% of the variance, therefore, the result of the relationship between total basic infrastructure investment and Gross Domestic Product can summarized by the linear model. 

Additionally, these results can confirm the expected finding is correct by the graph, the result of regression, and R-square.


# Additional Expected Findings

Question 18. It is expected that there will be an increase in almost all category's GDP except few which according to my assumption might have gone out of use so they are not really affecting GDP.

Question 20. It is expected that the chained GDP will increase the same way the current GDP is increasing because chained GDP is basically the index measure production in one year relative to another.

# Additional Analysis

*How each category is changed under Transportation from 2000 to 2017?*

```{r category}
change_transportation <- chain_investment %>%
  filter(meta_cat == "Transportation") %>%
  filter(between(year, 2000, 2017))

change_transportation %>%
  ggplot(aes(x= year,
             y= gross_inv_chain)) +
  geom_line() +
  facet_wrap(~ category) +
  ylab("Chained Gross Investment") +
  ggtitle("Change in transportation categories from 2000 to 2017")
```
\@ref(fig:category) shows the change in each category in the form of grid from year 2000 to 2017 under Transportation. It shows that some of the category remained constant with slow or little change in the chained gross investment but *Highway and streets* and *S&L* have shown a tremendous decrease being the highest invested category of GDP (large amount of money invested in it)


*What is the relationship between current GDP and chained GDP?*

```{r GDP_compare}
gdplev %>%
  ggplot(aes(x = GDP_current,
             y = GDP_chain)) + 
  geom_point(aes(color = year)) +
  geom_vline(aes(xintercept = mean(GDP_current))) +
  xlab("Real GDP") +
  ylab("Measure of GDP") +
  ggtitle("Relationship between Current and Chained GDP over the years")
```
\@ref(fig:GDP_compare) shows the yearly comparison between Real GDP and measure of GDP over the years. It is visible that in starting of the graph there is 0 GDP that is around 1920s. The horizontal line represents the average of Real GDP to see what is average GDP invested over the years and where the current GDP is lying (how far from the average). But I think the average is basically lined up because of 0 GDP in the starting of the years as most of the points lie there.

# Conclusion

This study explores infrastructural investment data in the USA and includes 4 interesting questions about the relationship between different categories of infrastructural investment and GDP. This study investigates the relationship between the total digital investment and GDP as a positive linear. Moreover, the correlative implication between transportation investment (chained US dollars) and GDP (chained US dollars) is a positive linear relationship. Furthermore, each association of private investment, federal investment, and GDP is a positive linear relationship and private investment is a higher effect on GDP than federal investment. In addition, total basic infrastructure investment has a positive linear relationship with GDP. In conclusion, the investment the study includes with GDP is all positive linear relationship, but different investments have different degrees of impact on GDP.


# References

Auguie B (2017). _gridExtra: Miscellaneous Functions for "Grid" Graphics_. R package
  version 2.3, <https://CRAN.R-project.org/package=gridExtra>.

Doménech, R., & Sicilia, J. (2021). Private investment as the engine of economic growth and social welfare / Private investment as the engine of economic growth and social welfare. https://www.bbvaresearch.com/wp-content/uploads/2021/04/EWPrivate_Investment_and_GDP_growthWB.pdf

Garrett Grolemund, Hadley Wickham (2011). Dates and Times Made Easy with lubridate.
  Journal of Statistical Software, 40(3), 1-25. URL https://www.jstatsoft.org/v40/i03/.

Gunnion, L. (2021, July). What does infrastructure investment mean for the US economy? Deloitte Insights. https://www2.deloitte.com/us/en/insights/economy/spotlight/impact-of-us-infrastructure-investment.html

H. Wickham. ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag New York, 2016.

IBM. (2022). About Linear Regression | IBM. Www.ibm.com. https://www.ibm.com/topics/linear-regression

Lin, X. (2020). Multiple pathways of transportation investment to promote economic growth in China: a structural equation modeling perspective. Transportation Letters, 12(7), 471–482. https://doi.org/10.1080/19427867.2019.1635780

Pebesma, E., 2018. Simple Features for R: Standardized Support for Spatial Vector Data.
  The R Journal 10 (1), 439-446, https://doi.org/10.32614/RJ-2018-009
  
Robinson D, Hayes A, Couch S (2022). _broom: Convert Statistical Objects into Tidy
  Tibbles_. R package version 1.0.0, <https://CRAN.R-project.org/package=broom>.  

Stupak, J. (2017). Economic Impact of Infrastructure Investment. https://ecommons.cornell.edu/bitstream/handle/1813/78289/CRS_Economic_Impact_of_Infrastructure_Investment_0717.pdf?sequence=1&isAllowed=y

Wickham H, Averick M, Bryan J, Chang W, McGowan LD, François R, Grolemund G, Hayes A,
  Henry L, Hester J, Kuhn M, Pedersen TL, Miller E, Bache SM, Müller K, Ooms J, Robinson
  D, Seidel DP, Spinu V, Takahashi K, Vaughan D, Wilke C, Woo K, Yutani H (2019). “Welcome
  to the tidyverse.” _Journal of Open Source Software_, *4*(43), 1686.
  doi:10.21105/joss.01686 <https://doi.org/10.21105/joss.01686>.
  
Wickham H, Bryan J (2022). _readxl: Read Excel Files_. R package version 1.4.0,
  <https://CRAN.R-project.org/package=readxl>.

Weisbrod, G., & Reno, A. (2009). ECONOMIC IMPACT OF PUBLIC TRANSPORTATION INVESTMENT. https://onlinepubs.trb.org/onlinepubs/tcrp/docs/TCRPJ-11Task7-FR.pdf

Yihui Xie (2022). knitr: A General-Purpose Package for Dynamic Report Generation in R. R
  package version 1.39.

Zhang, J., Zhao, W., Cheng, B., Li, A., Wang, Y., Yang, N., & Tian, Y. (2022). The Impact of Digital Economy on the Economic Growth and the Development Strategies in the post-COVID-19 Era: Evidence From Countries Along the “Belt and Road.” Frontiers in Public Health, 10. https://doi.org/10.3389/fpubh.2022.856142

Zhu H (2021). _kableExtra: Construct Complex Table with 'kable' and Pipe Syntax_. R
  package version 1.3.4, <https://CRAN.R-project.org/package=kableExtra>.


<details>
<summary>Session information</summary>
```{r, echo = FALSE}
sessioninfo::session_info()
```
</details>
