---
title: "ETC5521 Assignment 1"
subtitle: "Your title"
team: [lorikeet]
author:
  - [Aphiaut Imuan]
  - [Xintong You]
date: "`r Sys.Date()`"
output: 
  html_document
---

[This assignment is for ETC5521 Assignment 1 by Team `r rmarkdown::metadata$team` comprising of `r knitr::combine_words(rmarkdown::metadata$author)`.]{style="color:#006DAE;"}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE, 
                      warning = FALSE)
library(tidyverse)
library(ggplot2)
library(knitr)
library(lubridate)
library(sf)
library(kableExtra)
library(broom)
library(readxl)
library(gridExtra)
```

```{r, echo=FALSE}
investment <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-08-10/investment.csv')
chain_investment <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-08-10/chain_investment.csv')
ipd <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-08-10/ipd.csv')
gdplev <- read_excel("gdplev.xlsx", range = "A9:C101", 
    col_names = FALSE, col_types = c("numeric", 
        "numeric", "numeric"))
gdplev <- gdplev %>%
  rename("year" = ...1,
         "GDP_current" = ...2,
         "GDP_chain" = ...3)
```


# Introduction and motivation


Infrastructure investment is an important core of the country. It includes federal investments such as water suppliers and a private sector investment that manages electricity. It also includes the cost of research and development in technology.

Moreover, infrastructure investment plays a role in supporting both businesses and households. For example, the development of logistics has resulted in businesses reducing transportation costs. Households can access infrastructure services and can choose to use them. It also has a positive effect on both short-term and long-term economic growth [(Stupak, 2017)](https://ecommons.cornell.edu/bitstream/handle/1813/78289/CRS_Economic_Impact_of_Infrastructure_Investment_0717.pdf?sequence=1&isAllowed=y).

The above mentioned, the case study of infrastructure should be USA because the USA is a big country, has a large population, and people can access infrastructure services. Furthermore, the USA is also the world's largest economy. 

# Data description


The data comes from [Bureau of Economic Analysis](https://www.bea.gov/system/files/papers/BEA-WP2020-12.pdf). The raw .xlsx file is included, or can be downloaded directly from the [BEA Working paper series](https://www.bea.gov/system/files/2021-01/infrastructure-data-may-2020.xlsx).

The 3 primary data sets are already cleaned and saved as .csv files and they all have five variables. Of these, the four variables are the same as below.

```{r, echo=FALSE}
variable <- c("category","meta_cat","group_num","year")
class <- c("character","character","double","double")
description <- c("Category of investment",
                 "Group category of investment",
                 "Group number of investment",
                 "Year of investment")
e.g. <- c("Private,Federal...","Total infrastructure,Water supply,GDP...","1,2,3...","1947,1948,1949...")
dic<- data.frame(variable, class,description, e.g.)

kable(dic,
      align = "llll","pipe")
```

The 3 data sets record the gross investment, chained investment in millions of USD and Implicit Price Deflators (IPDs) respectively.

```{r, echo=FALSE}
variable <- c("gross_inv","gross_inv_chain","gross_inv_ipd")
class <- c("double","double","double")
description <- c("Gross investment in millions of USD",
                 "Gross investment (chained 2021 dollars) in millions of USD",
                 "Implicit Price Deflators (IPDs) An implicit price deflator is the ratio of the current-dollar value of a series, such as gross domestic product (GDP), to its corresponding chained-dollar value, multiplied by 100.")
dic2 <- data.frame(variable,class,description) 

kable(dic2,
      align = "lll","pipe")

```

The gdplev data set is downloaded from [Bureau of Economic Analysis](https://www.bea.gov/data/gdp/gross-domestic-product), which is a supplementary data set for our analysis. It records the current GDP and chain GDP in US from 1929 to 2021. Its variable contains year, GDP_current and GDP_chain whose type is all double.

- Chained dollars: Chained dollars is a method of adjusting real dollar amounts for inflation over time, to allow the comparison of figures from different years. The U.S. Department of Commerce introduced the chained-dollar measure in 1996.

# Questions of interest


1. What category of investment is the highest average gross investment and Chain? Does it same?

2. What is the trend of total basic infrastructure?

3. what is the trend of total social infrastructure?

4. What year and category are the highest investment and Chain?

5. What year and category are the lowest investment and Chain?

6. **What is the relationship between total digital investment and gdp?** 

7. What are the highest investment between Air, Water, Rail transportation investment ?

8. What are the similarities/difference of trend between Federal electric power structures investment and Private electric power structures investment?

9. How does the digital investment change? 

10. How does the transportation investment change? 

11. How does the power investment change? 

12. What is the proportion of total social investment in all category infrastructure?

13. What is the proportion of total basic infrastructure investment in all category infrastructure?

14. **What is the relationship between transportation investment (chaind us dollar) and gdp (chaind us dollar)?**

15. What are the trends in all categories infrastructure investment so far?

16. In 2012, what is the proportion of various investments in the total investment？

17. What category of investment has the greatest impact on GDP and how?

18. **What is the relationship between GDP and private investment, federal investment？And which one is stronger impact on GDP?**

19. What is the proportion of total digital investment in all category infrastructure?

20. **Is there a linear relationship between GDP and total basic infrastructure investment? Positive or negative?**

# Expected findings


In question 6: We expect the positive relationship between total data investment and GDP. Moreover we expect to see the steep line from 1900 to 2017.

In question 14: We expect the positive relationship between transportation (chaind US dollar) and GDP(chaind US dollar), the result can explain by the linear model and high effect value to GDP.

In question 18: We expect the strong relation between GDP and federal investment.

In question 20: We expect there is a positive linear relationship between GDP and total basic. infrastructure investment.

# Analysis and findings

**Methodology**
This study uses the linear model to analysis because linear models are the simple and present a mathematical equation which easy to interpret and can be predictions. Moreover, the linear regression model is a reliable predictors because it is a long established statistical procedure  and the character of linear regression is easy to understand (IBM, 2022).


**What is the relationship between total digital investment and gdp?**

```{r}
investment <- investment %>%
  left_join(gdplev, by = "year") 
investment <- investment %>%
  mutate(GDP_current = GDP_current*1000)


chain_investment <- chain_investment %>%
  left_join(gdplev, by = "year")

chain_investment <- chain_investment %>%
  mutate(GDP_current = GDP_current*1000)
```

```{r}
q6 <- investment %>%
  filter(category == "Total digital infrastructure") %>%
  rename("digital" = gross_inv)
  
q6_lm <-  lm(GDP_current ~ digital, data = q6)
  
ggplot(q6,
       aes(x = year,
           y = digital)) +
  geom_point() +
  geom_vline(xintercept = 1990, linetype="dashed", color = "red")+
  theme_bw() +
  ggtitle("The trend of total digital infrastructure investment from 1947 to 2017")


ggplot(q6,
       aes(x = year,
           y = GDP_current)) +
  geom_point() +
  geom_vline(xintercept = 1990, linetype="dashed", color = "red")+
  theme_bw() +
  ggtitle("The trend of Gross Domestic Product(GDP) from 1947 to 2017")

 ggplot(q6,
       aes(x = digital,
           y = GDP_current)) +
  geom_point(shape = 1,
               size = 2,
               color="#0072B2") +
  geom_smooth(method = "lm",
              col = "red",
              size = 1)+
  ggtitle("The relationship between total digital investment and gdp") +
   theme_bw()
q6_lm_summary <- glance(q6_lm)
q6_lm_summary
tidy(q6_lm)
```
The figure 1 shows the positive trend of total digital infrastructure investment, moreover, it is a steeply increasing trend between 1990 and 2017. Although overall the trend is increasing trend, around 1999 to 2009 are a slightly fluctuation.

The figure 3 illustrates the proportion increasing between Total digital infrastructure and GDP is 1:1. 
From the result can be expressed the formula as:

$$
\begin{align*}
{GDP} = 659595.97 + 112.63{digital}
\end{align*}
$$

This formula means if total digital infrastructure increase 1 million US dollar, the Gross Domestic Product will increase 112.63 million US dollar. This result related with study of Zhang et al. (2022) that investigated the increasing of digital economy will increase the GDP around 0.78%. Furthermore, they told it can be seen by the development of  new technology such as internet and mobile phone communication.

R-squared is 95.54% of variant and it shows nice linear model. Therefore, the result of relationship between Total digital infrastructure and GDP can summaries by linear model.

In Addition, this result same with our expected finding that is a positive relationship between total digital infrastructure and GDP. However, the trend of Total digital infrastructure has steeply increased since 1990, GDP has steeply increased since 1980. It means Total digital infrastructure doesn't significant affect to GDP from 1980 to 1990 too much.


**What is the relationship between transportation investment (chaind us dollar) and gdp (chaind us dollar)?**

```{r}

q14 <- chain_investment %>%
   filter(category == "Transportation") %>%
   pivot_wider(names_from = category, values_from = gross_inv_chain)

q14_lm <-  lm(GDP_chain ~ Transportation, data = q14)
  
transport <- ggplot(q14,
       aes(x = year,
           y = Transportation)) +
  geom_point() +
  theme_bw()#+
 # geom_vline(xintercept = 1990, linetype="dashed", color = "red")+
#  ggtitle("The trend of total digital infrastructure investment from 1947 to 2017")




gdp <- ggplot(q14,
       aes(x = year,
           y = GDP_chain)) +
  geom_point() +
  geom_vline(xintercept = 1990, linetype="dashed", color = "red")+
  theme_bw() +
  ggtitle("The trend of Gross Domestic Product(GDP) from 1947 to 2017")

grid.arrange(transport, gdp, nrow=2)

 ggplot(q14,
       aes(x = Transportation,
           y = GDP_chain)) +
  geom_point(shape = 1,
               size = 2,
               color="#0072B2") +
  geom_smooth(method = "lm",
              col = "red",
              size = 1)+
  ggtitle("The relationship between transportation investment and gdp") +
   theme_bw()
q14_lm_summary <- glance(q14_lm)
q14_lm_summary
tidy(q14_lm)
```

The figure 1 shows the positive trend of transportation investment (chained dollar), however, it is a steeply a slightly fluctuation between 1968 and 1998. While Gross Domestic Product (chained dollar) is a positive trend and increases over year.

The figure 3 illustrates the relationship between transportation investment (chained dollar) and GDP (chained dollar),  nonetheless it is difficult explain the relationship of both variables is linear. 
From the result can be expressed the formula as:

$$
\begin{align*}
{GDP} = -5164.83 + 0.13{Transportation}
\end{align*}
$$

This formula means if transportation investment increase 1 million chained US dollar, the Gross Domestic Product will increase 0.13 million chained US dollar. This result relates with the study of Weisbrod and Reno (2009) that agrued the positive significant of relation between transportation investment and GDP. Moreover, transportation investment has a positive effect to economic	growth (Lin, 2020). 

R-squared is 67.39% of variant Therefore, the result of relationship between transportation investment (chained dollar) and Gross Domestic Product (chained dollar) should not summary by linear model or this formula has omitted variables. 

This result same with our expected finding that is a positive relationship between transportation investment (chained dollar) and Gross Domestic Product (chained dollar). However, this relationship can not explain with linear model and the coefficient is too low that are not same with our expected finding.

**What is the relationship between GDP and private investment, federal investment？And which one is stronger impact on GDP?**

```{r q18}
options(scipen=100, digits=4)
q18_P1 <- investment %>%
  filter(category == "Private") 
q18_P <- q18_P1 %>%
  filter(meta_cat == "Total infrastructure")
q18_F1 <- investment %>%
  filter(category == "Federal")
q18_F <- q18_F1 %>%
  filter(meta_cat == "Total infrastructure")

gg1 <- ggplot(q18_P,aes(year,GDP_current)) +
  geom_line(color = "blue") +
  theme_bw() +
  ggtitle("Trend of current GDP") 

gg2 <- ggplot(q18_P,aes(year,gross_inv)) +
  geom_line(color="blue") +
  theme_bw()+
  ggtitle("Trend of gross private investment") 

gg3 <-  ggplot(q18_F,aes(year,gross_inv)) +
  geom_line(color="blue") +
  theme_bw() +
  ggtitle("Trend of gross federal investment") 
  
grid.arrange(gg1,gg2,gg3,layout_matrix=rbind(c(1,1),c(2,3)))

ggplotRegression <- function (x) {
  ggplot(x$model, aes_string(x = names(x$model)[2], 
                               y = names(x$model)[1])) +
    geom_point(shape = 1,
               size = 2,
               color="#0072B2")+
    geom_smooth(method = "lm", 
                col = "red",
                size = 1) +
    labs(title = paste("R2 = ",signif(summary(x)$r.squared, 3)))
}

g1 <- ggplotRegression(lm(GDP_current ~ gross_inv,data = q18_P)) +
  labs(x = "Gross Private Investment", y = "Current GDP") +
  theme_bw()
g2 <- ggplotRegression(lm(GDP_current ~ gross_inv,data = q18_F)) +
  labs(x = "Gross Federal Investment", y = "Current GDP") +
  theme_bw()

grid.arrange(g1,g2,nrow = 1,top = "Linear relationship between investment and GDP")
```

As we can see in the line graph, the current GDP has a generally increasing trend by time, and the trend of gross private investment is similar with that. While the gross federal investment has fluctuated since around 1980.

By observing the trends of the three variables, we guess that current GDP and the other two may have a linear relationship, so we make a linear model to judge whether they have a linear relationship. As a result, GDP dose have linear relationship with private investment and federal investment respectively, and the private investment has stronger impact on current GDP.

R-squared (R2) is a important statistical measurement that represents the proportion of the variance for current GDP that’s explained by the gross private and federal investment in the regression model. For example, 98.6% private investment can explain the current GDP, rather only 78.1% federal investment can explain that. Obviously, there is a stronger linear relationship between GDP and private investment which is unexpected. This result relates with the study of Private investment as the engine of economic growth and social welfare that investigated the private investment can explain the variance of GDP growth rate more than public investment (Doménech & Sicilia, 2021).

**Is there a linear relationship between GDP and total basic infrastructure investment? Positive or negative?**

```{r}
q20 <- investment %>%
  filter(category == "Total basic infrastructure")

q20_lm <-lm(GDP_current ~ gross_inv, data = q20)
q20_lm_summary <- glance(q20_lm)
q20_lm_summary
tidy(q20_lm)

g3 <- ggplotRegression(lm(GDP_current ~ gross_inv,data = q20)) +
  labs(x = "Total Basic Infrastructure Inv", y = "Current GDP") +
  theme_bw() +
  ggtitle("Relationship between GDP and total basic infrastructure investment")
g3
```

From the result can be expressed the formula as:

$$
\begin{align*}
{GDP} = -337073.45	 + 59.02 {Basic}
\end{align*}
$$

This formula means if transportation investment increase 1 million US dollar, the Gross Domestic Product will increase 59.02  million US dollar. Moreover, the distribution of data has a positive linear relationship in figure 4. The trend is increasing because infrastructure investment effect to increasing a huge stock of public capital such as developing or creating new roads. That is a directly effect to increase government spending(G) in the gdp formula (Stupak, 2017).  Furethermore, Gunnion (2021) claims the investment of infrastructure is positive impact to economic output.

R-squared is 98.67% of variant, therefore, the result of relationship between total basic infrastructure investment and Gross Domestic Product can summary by linear model. 

Additionally, these result can confirm the expected finding is correct by the graph, the result of regression and R-square.


# References

Doménech, R., & Sicilia, J. (2021). Private investment as the engine of economic growth and social welfare / Private investment as the engine of economic growth and social welfare. https://www.bbvaresearch.com/wp-content/uploads/2021/04/EWPrivate_Investment_and_GDP_growthWB.pdf
Gunnion, L. (2021, July). What does infrastructure investment mean for the US economy? Deloitte Insights. https://www2.deloitte.com/us/en/insights/economy/spotlight/impact-of-us-infrastructure-investment.html
IBM. (2022). About Linear Regression | IBM. Www.ibm.com. https://www.ibm.com/topics/linear-regression
Lin, X. (2020). Multiple pathways of transportation investment to promote economic growth in China: a structural equation modeling perspective. Transportation Letters, 12(7), 471–482. https://doi.org/10.1080/19427867.2019.1635780
Stupak, J. (2017). Economic Impact of Infrastructure Investment. https://ecommons.cornell.edu/bitstream/handle/1813/78289/CRS_Economic_Impact_of_Infrastructure_Investment_0717.pdf?sequence=1&isAllowed=y
Weisbrod, G., & Reno, A. (2009). ECONOMIC IMPACT OF PUBLIC TRANSPORTATION INVESTMENT. https://onlinepubs.trb.org/onlinepubs/tcrp/docs/TCRPJ-11Task7-FR.pdf
Zhang, J., Zhao, W., Cheng, B., Li, A., Wang, Y., Yang, N., & Tian, Y. (2022). The Impact of Digital Economy on the Economic Growth and the Development Strategies in the post-COVID-19 Era: Evidence From Countries Along the “Belt and Road.” Frontiers in Public Health, 10. https://doi.org/10.3389/fpubh.2022.856142

